The origin of cosmic-ray electrons in cluster outskirts

### 2011-11-23 paper.tex
updated text

% --- Anders's commands --- %
\newcommand{\ev}{\mathrm{eV}}
\newcommand{\gev}{\mathrm{GeV}}
\newcommand{\mev}{\mathrm{MeV}}
\newcommand{\tev}{\mathrm{TeV}}
\newcommand{\mpc}{m_{\mathrm{p}}c}
\newcommand{\mpcc}{m_{\mathrm{p}}c^2}
\newcommand{\p}{\mathrm{p}}
\newcommand{\pp}{\mathrm{pp}}
\newcommand{\br}{\bold{r}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\inj}{\mathrm{inj}}
\newcommand{\pd}{\partial}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\gas}{\mathrm{gas}}
\newcommand{\mr}{\mathrm}
\newcommand{\DM}{\mathrm{DM}}
\newcommand{\ISM}{\mathrm{ISM}}
\newcommand{\ICM}{\mathrm{ICM}}
\newcommand{\CMB}{\mathrm{CMB}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\mecc}{m_\e c^2}
\newcommand{\mug}{\mathrm{\mu G}}
\newcommand{\rvir}{R_{200}}
\newcommand{\B}{\mathrm{B}}
\newcommand{\ma}{\mathrm{max}}
\newcommand{\low}{\mathrm{low}}
\newcommand{\gae}{\gamma_\mathrm{e}}

@@

\subsection{CR electron distribution function} 
The injected CR electron distribution function at each time $t$ is
given by
\begin{eqnarray}
  \label{eq:f_inj_e}
  f_\rmn{inj,e}(p_\e) &=&  \frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd p_\e}(p_\e) = 
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd p_\p}(p_\e)\frac{m_\e}{m_\p} \nonumber\\
&=& \frac{\dd^2 N_\rmn{inj,p}}{\dd V \dd p_\p}(p_\p)
\left(\frac{p_\e}{p_\p}\right)^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}\frac{m_\e}{m_\p} = 
\Delta C\,p_\e^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}\frac{m_\e}{m_\p}\,.
 \nonumber\\
&&
\end{eqnarray}
and $p_\e = P_\e/m_\e c$, where we have normalized the electron
momentum $P_\e$ with the electron mass $m_\p$. We use that maximal
50\% of the energy available in a shock is injected into CR protons
($\eta_\rmn{max,p}$) and a factor 10 smaller efficiency of 5\% for the
CR electrons ($\eta_\rmn{max,e}$).

The CR electrons cool through inverse Compton (IC) emission and
Coulomb losses on timescales that are relative short compared to the
dynamical timescale of a cluster. We model these looses analytically
by assuming a self similar solution for the CR electron distribution
function \citep{1999ApJ...520..529S}
\begin{eqnarray}
  \label{eq:f_e}
  f_\e(\gamma_\e) &\approx& \frac{f_\rmn{inj,e}(\gamma_\e)}{\alpha-1}
\left(\frac{\gamma_\low}{\gamma_\e}+\frac{\gamma_\e }{\gamma_\ma}\right)^{-1}
\nonumber\\
&\times& \left\{2-\left[\frac{1}{2}\left(1-\frac{\gamma_\e}{\gamma_\ma}+
\left|1-\frac{\gamma_\e}{\gamma_\ma}\right|\right)\right]^{\alpha-1}\right.
\nonumber\\
&-& \left.\left(1+\frac{\gamma_\low}{\gamma_\e}\right)^{-(\alpha-1)}\right\}\,,
\end{eqnarray}
Here $\gamma_\e \approx p_\e = P_\e/m_\e c$ for $ p_\e \gg 1$.  The
approximation in Eqn.~(\ref{eq:f_e}) is valid as long as
$\gamma_\ma\gg \gamma_\low$, where the upper limit is determined from
IC losses through
\begin{eqnarray}
  \label{eq:gamma_max}
b_\IC(\gamma_\e) &=& b_\IC \gamma_\e^2 
= \frac{4}{3}\frac{\sigma_\rmn{T}}{m_\e c} U_\rmn{cmb} \gamma_\e^2 \,,\quad \rmn{and}\nonumber \\
%1.37\times 10^{-20} \gamma^2(1+z)^4 \,\rmn{s}^{-1}
\gamma_\ma &=& \frac{1}{b_\IC\,\Delta t}\,.
\end{eqnarray}
Here $\sigma_\rmn{T}= 8\pi e^4/3(m_\e c^2)^2$ is the Thomson cross
section and $U_\rmn{cmb}$ is the energy density of the CMB in the
cluster. The lower limit is determined from Coulomb losses through
\begin{eqnarray}
  \label{eq:gamma_min}
b_\rmn{C}(\gamma_\e) &=& b_\rmn{C} \gamma_\e^2 
= \frac{3\,\sigma_\rmn{T}\,n_\rmn{el}\,c}{2}
\left[\ln\left(\frac{m_\e c^2 \sqrt{\gamma_\e-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
  &-&\ln(2)\left(\frac{1}{2}+\frac{1}{\gamma_\e}\right)+\frac{1}{2}+
\left.\left(\frac{\gamma_\e-1}{4\gamma_\e}\right)^2\right]
         \,,\quad \rmn{and}\nonumber \\
\gamma_\low &=& b_\rmn{C}\,\Delta t\,.
\end{eqnarray}
Here $\omega_\rmn{plasma} = \sqrt{4\pi e^2 n_\e / m_\e}$ is the plasma
frequency, and $n_\e$ is the number density of free electrons.

We model the final CR electron spectrum as a superposition of five CR
populations, each determined by Eqn.~\ref{eq:f_e} but with a
different spectral index and injected spectra.

\subsection{Synchrotron emission}
WORK UNDER PROGRESS During the merging of two clusters, both
turbulence flows and merging shocks are induced. These processes are
thought to accelerate CRs to high energy, where especially the
distribution function of preexisting CRs is enhanced. In this section
we use the smooth CR electron distribution derived from a simulated
merging cluster in the previous section, boost it using the details of
the shock, and then finally derive the radio synchrotron emission.

We fit the steady state CR electron distribution function in
Fig.~\ref{fig:e_spec} with two power-laws
\begin{equation}
  f_\rmn{e,fit}(\gamma_\e) = C_1\,\gamma_\e^{-\alpha_1}
\theta(\gamma_\rmn{break}-\gamma_\e) 
+ C_2\,\gamma_\e^{-\alpha_2}
\theta(\gamma_\e-\gamma_\rmn{break})\,. 
\label{eq:f_fit}
\end{equation}
The best fit CR electron normalization is given by ${\bf
  C}=(6\times10^{-4},\,10^{7})$, the spectral index by ${\bf
  \alpha}=(3.0,5.5)$, and the energy where the slope changes by
$\gamma_\rmn{break}=1.2\times10^4$.

\C{Probably need to calculate the emissivity numerically since there
  is a break in the power-law spectrum. CHECK} We calculate the radio
synchrotron emissivity of the CR electrons using \cite{1979rpa..book.....R}
\begin{equation}
J_0(\nu) \approx \left(\frac{e^2\,\nu_c}{c}\right)\,\sum_i C_i
\Gamma\left(\frac{3 \alpha_i-1}{12}\right)\,\Gamma\left(\frac{3 \alpha_i+19}{12}\right)\frac{3^\frac{\alpha_i}{2}}{\alpha_i+1}\,\left(\frac{\nu}{\nu_c}\right)^{\frac{1-\alpha}{2}}\,,
\end{equation}
where the cyclotron frequency is given by
$\nu_c=e\,B\,c/(2\pi\mecc)$. We derive the total emitted power from
the source by integrating the emissivity over the target volume. Here,
we assume that the target volume haw an uniform distribution of the
CR electrons, hence the total power per unit frequency is given by
\begin{equation}
  P_0(\nu) = \rmn{Volume} \times J_0(\nu)\,.
\end{equation}

@@

The steady-state test-particle solution of the downstream CR
distribution can be written as a function of the pre-existing CRs through
\begin{equation}
  f_2(\gae) = q \gae^{-q}\int_{\gamma_\rmn{inj}}^\gamma \gamma'^{q-1}
   f_\rmn{e,fit}(\gamma')\dd \gamma'\,.
\end{equation}
Here we have neglected the contribution from the injected CRs. The lowest momentum boundary above which particles can cross the shock is given by 
\begin{equation}
\gamma_\rmn{inj} \approx 1.17 \frac{u_2}{c}\left(1+\frac{1.07}{\epsilon_B}\right)
\left(\frac{M}{3}\right)^{0.1}\,.
\end{equation}

Spectral index
\begin{equation}
q(M) = \frac{3\left(u_1-v_ A\right)}{u_1-v_A-u_2} = 
\frac{3\sigma\left(1-M_A^{-1}\right)}{\sigma-1-\sigma M_A^{-1}}
\end{equation}
where $u_1$ and $u_2$ are the speed of the flow upstream and
downstream, respectively, in the shock rest frame. The compression
ratio is given by $\sigma=u_1/u_2=\rho_2/\rho_1$, and $M_A=u_1/v_A$ is
the upstream Alfvén Mach number with velocity
$v_A=B_1/\sqrt{4\pi\rho_1}$. We derive the test-particle power-law
slope $q$ as a function of the shock Mach number $M$ with
$\sigma=[(\gamma_\rmn{ad}+1)M^2]/[(\gamma_\rmn{ad}-1)M^2+2]$, where
adiabatic index $\gamma_\rmn{ad}=5/3$, $M_A=M c_s/v_A$, and $ c_s$ the
upstream sound speed.

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/CRespec0.3Rv.140.v12.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/CRespec1.4Rv.140.v12.eps}
  \caption{Cosmic ray electron spectra in cluster outskirts at
    redshift zero. We show the CR electron distribution function
    weighted with the electron Lorentz factor ($\gamma_\e$)
    squared. In the {\it left panel} panel we show the CR electron in
    the region between (0.3-0.5)$\rvir$ and in the {\it right panel}
    between (1.3-1.5)$\rvir$. The solid lines show the median CR
    electron spectrum where both radiative and adiabatic losses are
    included. The dotted lines shows the 68 percentiles. The dashed
    lines show the CR electron spectrum without Coulomb and radiative
    cooling. Note that the Coulomb cooling at low energies is very
    inefficient because of the low electron densities in the cluster
    outskirts. \label{fig:e_spec}}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/f_evolution.0.3Rv.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/f_evolution.1.4Rv.eps}
  \caption{Time evolution of cosmic ray (CR) electron spectra. We show
    the CR electron distribution function weighted with the electron
    Lorentz factor ($\gamma_\e$) squared. The time evolution of the CR
    electrons which at $z=0$ reside from (0.3-0.5)$\rvir$ are shown in
    the {\it left panel} and (1.3-1.5)$\rvir$ in the {\it right
      panel}. The black lines show the CR electrons with a
    $\gamma_\e=30$, blue lines $\gamma_\e=200$, and green lines
    $\gamma_\e=10^4$. The solid lines show the CR electrons with full
    cooling, i.e. for Coulomb, inverse Compton, and adiabatic
    losses. Dashed lines show the CR electrons without Coulomb and
    inverse Compton cooling while the dash-dotted lines show the CR
    electrons without adiabatic cooling. The dotted line show the
    injected CR electron distribution function, derived from the
    injected CR protons. For each energy, we normalize the CR electron
    distribution function with the CR electrons without
    cooling. Notice that the high energy part is built up during the
    last Gyr. This is explained by the inverse Compton losses that are
    larger at high energies, especially at high redshifts where the
    energy density of the CMB is much larger than
    today.  \label{fig:e_evol}}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/evolution.0.3Rv.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/evolution.1.4Rv.eps}
  \caption{Time evolution of the gas density and the cosmic ray (CR)
    cooling time. We show the time evolution of the electron number
    density and CR electron cooling time that are associated with the
    CR electrons which at redshift $z=0$ reside from (0.3-0.5)$\rvir$
    in the {\it left panel} and (1.3-1.5)$\rvir$ in the {\it right
      panel}. The red lines show the electron number densities, where
    the dashed line show the median and dash-dotted the mean. The
    solid lines show the cooling times
    $\tau_\rmn{cool}=\gamma_\e/[b_\IC(\gamma_\e)+b_\rmn{C}(\gamma_\e)]$
    for $\gamma_\e=30$ (black), $\gamma_\e=200$ (blue), and
    $\gamma_\e=10^4$ (green). Interesting the CR cooling at low
    energies of the particles that end up in the center is constant
    while is increasing monotonically for the particles that end up in
    the cluster periphery. The density of the particles that end up in
    the center is roughly constant with time while it is decreasing
    for the particles that end up in the outer part. This is due to
    the expansion of the universe that counteract the contraction of
    the particles in the cluster periphery. \label{fig:evol}}
\end{minipage}
\end{figure*}

>>>>

% --- Anders's commands --- %
\newcommand{\ev}{\mathrm{eV}}
\newcommand{\gev}{\mathrm{GeV}}
\newcommand{\mev}{\mathrm{MeV}}
\newcommand{\tev}{\mathrm{TeV}}
\newcommand{\mpc}{m_{\mathrm{p}}c}
\newcommand{\mpcc}{m_{\mathrm{p}}c^2}
\newcommand{\p}{\mathrm{p}}
\newcommand{\Pp}{\mathrm{P}_\mathrm{p}}
\newcommand{\Pe}{\mathrm{P}_\mathrm{e}}
\newcommand{\pp}{p_\mathrm{p}}
\newcommand{\pe}{p_\mathrm{e}}
\newcommand{\br}{\bold{r}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\inj}{\mathrm{inj}}
\newcommand{\pd}{\partial}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\gas}{\mathrm{gas}}
\newcommand{\mr}{\mathrm}
\newcommand{\DM}{\mathrm{DM}}
\newcommand{\ISM}{\mathrm{ISM}}
\newcommand{\ICM}{\mathrm{ICM}}
\newcommand{\CMB}{\mathrm{CMB}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\mecc}{m_\e c^2}
\newcommand{\mug}{\mathrm{\mu G}}
\newcommand{\rvir}{R_{200}}
\newcommand{\B}{\mathrm{B}}
\newcommand{\ma}{\mathrm{max}}
\newcommand{\low}{\mathrm{low}}
\newcommand{\gae}{\gamma_\mathrm{e}}
\newcommand{\gam}{\gamma}

@@

\subsection{CR electron distribution function} 
In a hot plasma the temperatures of electrons and protons (ions) are
equal. Under this constraint we derive the relationship between CR
electon momentum and CR proton momentum given by
%\begin{eqnarray}
%  \label{eq:pe_pp}
%\dd \Pe = \dd \Pp \frac{\sqrt{\Pe^2+2\left(m_\e c\right)^2-2m_\e m_\p\,c^2
%+2\left(m_\p-m_\e\right)c\sqrt{\left(m_\e c\right)^2+\Pe^2}}}{\Pe+\left(m_\p-m_\e\right) c \frac{\Pe}{\sqrt{\left(m_\e c\right)^2+\Pe^2}}}
%\end{eqnarray}
\begin{eqnarray}
  \label{eq:pe_pp}
\dd \pe &=& \dd \pp \,g(\pe) =  \dd \pp \,h(\pp)\,,\qquad \rmn{where}\\
g(\pe) &=& X \,\frac{\left[\pe^2-2\left(X-1\right)\left(1-\sqrt{1+\pe^2}\right)\right]^{0.5}}{\displaystyle\pe+\left(X-1\right)\frac{\pe}{\sqrt{1+\pe^2}}}\,,\qquad \rmn{and}\\
h(\pp) &=& X \,\frac{\displaystyle\pp-\left(1-Y\right)\frac{\pp}{\sqrt{1+\pp^2}}}{\left[\pp^2+2\left(1-Y\right)\left(1-\sqrt{1+\pp^2}\right)\right]^{0.5}}\,.\\
\end{eqnarray}
Note that $\pe = \Pe/m_\e c$ and $\pp = \Pp/m_\p c$, where we have
normalized the electron momentum $\Pe$ and the proton momentum $\Pp$
with the electron mass $m_\e$ and proton mass $m_\p$, respectively. We
have introduced the mass ratios $X=m_\p/m_\e$ and $Y=m_\e/m_\p$. Also
note that $\dd \Pe\rightarrow \dd \Pp$ in the asymptotic limit where
$\pp\gg 1$, and that $\dd \Pe\rightarrow \sqrt{m_\e / m_\p}\,\dd \Pp$ for
$\pp\ll 1$.

The injected CR electron distribution function at each time $t$ is
given by
\begin{eqnarray}
  \label{eq:f_inj_e}
  f_\rmn{inj,e}(p_\e) &=&  \frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd p_\e}(p_\e) = 
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd p_\p}(p_\e)\frac{1}{g(p_\e)} \nonumber\\
&=& \frac{\dd^2 N_\rmn{inj,p}}{\dd V \dd p_\p}(p_\p)
\left(\frac{p_\e}{p_\p}\right)^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}\frac{1}{g(p_\e)} = 
\Delta C\,p_\e^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}\frac{1}{g(p_\e)}\,.
 \nonumber\\
&&
\end{eqnarray}
We use that maximal 50\% of the energy available in a shock is
injected into CR protons ($\eta_\rmn{max,p}$) and a factor 10 smaller
efficiency of 5\% for the CR electrons ($\eta_\rmn{max,e}$).

The CR electrons cool through inverse Compton (IC) emission and
Coulomb losses on timescales that are relative short compared to the
dynamical timescale of a cluster. We model these looses analytically
by instantiniuosly injecting a power-law of electrons at time $t_i$
and evolving it to a later time $t$ (for further details see
\citep{1999ApJ...520..529S}). The loss of energy for each particle is
described by
\begin{equation}
\label{eq:d_cool}
\frac{\dd \gam}{\dd t} = - b(\gam,t)\,,
\end{equation}
where the loss function $b(\gam,t)$ is dominated by Coulomb and IC
losses for the energies and gas densities relevant in this work. The
Coulomb losses are given by
\begin{eqnarray}
  \label{eq:b_C}
  b_\rmn{C}(\gam) &=& b_\rmn{C} \gam^2 
  = \frac{3\,\sigma_\rmn{T}\,n_\rmn{el}\,c}{2}
  \left[\ln\left(\frac{m_\e c^2 \sqrt{\gam-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{1}{2}+\frac{1}{\gam}\right)+\frac{1}{2}+
    \left.\left(\frac{\gam-1}{4\gam}\right)^2\right]\,.
\end{eqnarray}
Here $\omega_\rmn{plasma} = \sqrt{4\pi e^2 n_\e / m_\e}$ is the plasma
frequency, and $n_\e$ is the number density of free electrons. Since
the Coulomb losses for relativistic electrons is almost independent of
energy (logarithmic), we approximate that $b_\rmn{C}(\gam,t)\approx
b_\rmn{C}(t)$ \footnote{We use a constant energy $\gamma=10^2$ to
  model the very weak dependence of energy, and note that the
  particular value of $\gamma$ does not matter as long as
  $\gamma\gg1$.}. We can now derive the shift in energy $\gam_i$ at
time $t_i$ to energy $\gam$ at time $t$ due to Coulomb cooling by
integrating Eqn.~(\ref{eq:d_cool}):
\begin{eqnarray}
  \label{eq:gamma_low}
  \gam_\low \equiv\int_{t_i}^{t}\dd t b_\rmn{C}(t) = 
  -\int_{\gam_i}^{\gam}\dd \gam' = -(\gam-\gam_i)\,.
\end{eqnarray}
Since we have discrete snapshots in time we can approximate 
\begin{eqnarray}
  \label{eq:gamma_low_sum}
  \gam_\low = \Delta t \sum_j b_\rmn{C}(t_j)\,,
\end{eqnarray}
where the sum includes all snapshots from time $t_i$ to time $t$. We
now continue with the IC losses that are given by
\begin{equation}
  \label{eq:b_IC}
  b_\IC(\gam,z) = b_\IC \gam^2 (1+z)^4
  = \frac{4}{3}\frac{\sigma_\rmn{T}}{m_\e c} U_\rmn{cmb} \gam^2\,.
\end{equation}
Here $\sigma_\rmn{T}= 8\pi e^4/3(m_\e c^2)^2$ is the Thomson cross
section and $U_\rmn{cmb}$ is the energy density of the CMB at redshift
$z=0$. Similarly to the Coulomb cooling we derive the evolution of energy 
of a particle subject to IC losses through
\begin{equation}
  \label{eq:b_IC_evolu}
\frac{\dd \gam}{\gam^2} = -b_\IC (1+z)^4\dd t\,.
\end{equation}
After a time $(t-t_i)$ has elapsed, all the electrons with
energies above $\gam_\ma$ have thermalized, where we derive $\gam_\ma$
through integrating Eqn.~(\ref{eq:b_IC_evolu}) from injected time
$z_i=z(t_i)$ to a later time $z=z(t)$:
\begin{equation}
  \label{eq:b_max}
   \frac{1}{\gam_\ma} \equiv \frac{1}{\gam} - \frac{1}{\gam_i} = 
   \frac{b_\IC}{H_0}\left[\Lambda(z)-\Lambda(z_i)\right]\,.
\end{equation}
Here $\Lambda(z) \approx z + 1.23\,z^2 + 0.50\,z^3 - 0.14\,z^4 -
0.04\,z^5$ in a Lambda CDM universe and
$H_0=h\,\rmn{km}\,\rmn{s}^{-1}\,\rmn{Mpc}^{-1}$ is the Hubble constant.

Given an initial energy $\gam_i$ of an electron at time $t_i$,
Eqn.~(\ref{eq:d_cool}) can be integrated to give the value of $\gam$
at a later time $t$. The differential population density for
relativistic electrons ($\gam_\e \approx p_\e = P_\e/m_\e c$) is then
given by
\begin{eqnarray}
\label{eq:f_evolv}
f_\rmn{inj,e}(\gam,t,t_i) &=& f_\rmn{inj,e}(\gam_i,t_i)
\left.\frac{\partial\gam_i}{\partial\gam}\right|_t\,,\quad \rmn{where} \\
 f_\rmn{inj,e}(\gam_i,t_i) &=& 
f_\rmn{inj,e}(\gam-\Delta\gam_\IC-\Delta\gam_\rmn{C},t_i) \nonumber\\
&=& f_\rmn{inj,e}(\gam+\frac{\gam^2}{\gam_\ma-\gam}+\gam_\low,t_i)\,,\quad \rmn{and} \\
\frac{\partial\gam_i}{\partial\gam} &=& 
\frac{\gam_\ma^2}{\left(\gam_\ma-\gam\right)^2}\,.
\end{eqnarray}
Here we have used that $\Delta\gam_\IC=-\gam_\low$ and
$\Delta\gam_\rmn{C}=\frac{-\gam^2}{\gam_\ma-\gam}$. The total electron
spectrum is derived from the sum of all individually cooled injected
spectra, starting from the time of injection $t_i$ until a later time
$t$,
\begin{eqnarray}
\label{eq:f_sum}
f_\rmn{tot,e}(\gam,t) = \left(\frac{\tilde{\rho}_{gas}(t)}{\rho_0}\right)^{-1/3}\sum_j f_\rmn{inj,e}(\gam,t,t_j)\,.
\end{eqnarray}
Here the comoving gas density factor $\rho_{gas}(t)$ takes care of the
adiabatic gains and losses, where $\rho_0=1$ is a reference density in
GADGET. Also note that the electrons injected at time $t=t_i$ have not
had the time to cool yet.

We model the final CR electron spectrum as a superposition of five CR
populations, each determined by Eqn.~\ref{eq:f_sum} but with a
different spectral index and injected spectra.

\subsection{Synchrotron emission}
The process of merging clusters is often very violent where large
amounts of gravitational energy beeing dissipated in the form of
radiation, increased temperature, turbulence flows and shocks. The two
latter procesess are belived to accelerate CRs to high energy, where
especially the distribution function of preexisting CRs is
enhanced. In this section we use the smooth CR electron distribution
derived from a simulated merging cluster in the previous section,
boost it using the details of the shock, and then finally derive the
radio synchrotron emission.

We fit the steady state CR electron distribution function in
Fig.~\ref{fig:e_spec} with two power-laws
\begin{equation}
  f_\rmn{e,fit}(\gam_\e) = C_1\,\gam_\e^{-\alpha_1}
\theta(\gam_\rmn{break}-\gam_\e) 
+ C_2\,\gam_\e^{-\alpha_2}
\theta(\gam_\e-\gam_\rmn{break})\,. 
\label{eq:f_fit}
\end{equation}
The best fit CR electron normalization is given by ${\bf
  C}=(6\times10^{-4},\,10^{7})$, the spectral index by ${\bf
  \alpha}=(3.0,5.5)$, and the energy where the slope changes by
$\gam_\rmn{break}=1.2\times10^4$.

\C{Probably need to calculate the emissivity numerically since there
  is a break in the power-law spectrum. CHECK} We calculate the radio
synchrotron emissivity of the CR electrons using \cite{1979rpa..book.....R}
\begin{equation}
J_0(\nu) \approx \left(\frac{e^2\,\nu_c}{c}\right)\,\sum_i C_i
\gam\left(\frac{3 \alpha_i-1}{12}\right)\,\gam\left(\frac{3 \alpha_i+19}{12}\right)\frac{3^\frac{\alpha_i}{2}}{\alpha_i+1}\,\left(\frac{\nu}{\nu_c}\right)^{\frac{1-\alpha}{2}}\,,
\end{equation}
where the cyclotron frequency is given by
$\nu_c=e\,B\,c/(2\pi\mecc)$. We derive the total emitted power from
the source by integrating the emissivity over the target volume. Here,
we assume that the target volume haw an uniform distribution of the
CR electrons, hence the total power per unit frequency is given by
\begin{equation}
  P_0(\nu) = \rmn{Volume} \times J_0(\nu)\,.
\end{equation}

@@

The steady-state test-particle solution of the downstream CR
distribution can be written as a function of the pre-existing CRs through
\begin{equation}
  f_2(\gae) = q \gae^{-q}\int_{\gam_\rmn{inj}}^\gam \gam'^{q-1}
   f_\rmn{e,fit}(\gam')\dd \gam'\,.
\end{equation}
Here we have neglected the contribution from the injected CRs. The lowest momentum boundary above which particles can cross the shock is given by 
\begin{equation}
\gam_\rmn{inj} \approx 1.17 \frac{u_2}{c}\left(1+\frac{1.07}{\epsilon_B}\right)
\left(\frac{M}{3}\right)^{0.1}\,.
\end{equation}

Spectral index
\begin{equation}
q(M) = \frac{3\left(u_1-v_ A\right)}{u_1-v_A-u_2} = 
\frac{3\sigma\left(1-M_A^{-1}\right)}{\sigma-1-\sigma M_A^{-1}}
\end{equation}
where $u_1$ and $u_2$ are the speed of the flow upstream and
downstream, respectively, in the shock rest frame. The compression
ratio is given by $\sigma=u_1/u_2=\rho_2/\rho_1$, and $M_A=u_1/v_A$ is
the upstream Alfvén Mach number with velocity
$v_A=B_1/\sqrt{4\pi\rho_1}$. We derive the test-particle power-law
slope $q$ as a function of the shock Mach number $M$ with
$\sigma=[(\gam_\rmn{ad}+1)M^2]/[(\gam_\rmn{ad}-1)M^2+2]$, where
adiabatic index $\gam_\rmn{ad}=5/3$, $M_A=M c_s/v_A$, and $ c_s$ the
upstream sound speed.

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
%  \includegraphics[width=0.49\columnwidth]{./figures/CRespec0.3Rv.140.v12.eps}
%  \includegraphics[width=0.49\columnwidth]{./figures/CRespec1.4Rv.140.v12.eps}
  \caption{Cosmic ray electron spectra in cluster outskirts at
    redshift zero. We show the CR electron distribution function
    weighted with the electron Lorentz factor ($\gam_\e$)
    squared. In the {\it left panel} panel we show the CR electron in
    the region between (0.3-0.5)$\rvir$ and in the {\it right panel}
    between (1.3-1.5)$\rvir$. The solid lines show the median CR
    electron spectrum where both radiative and adiabatic losses are
    included. The dotted lines shows the 68 percentiles. The dashed
    lines show the CR electron spectrum without Coulomb and radiative
    cooling. Note that the Coulomb cooling at low energies is very
    inefficient because of the low electron densities in the cluster
    outskirts. \label{fig:e_spec}}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
%  \includegraphics[width=0.49\columnwidth]{./figures/f_evolution.0.3Rv.eps}
%  \includegraphics[width=0.49\columnwidth]{./figures/f_evolution.1.4Rv.eps}
  \caption{Time evolution of cosmic ray (CR) electron spectra. We show
    the CR electron distribution function weighted with the electron
    Lorentz factor ($\gam_\e$) squared. The time evolution of the CR
    electrons which at $z=0$ reside from (0.3-0.5)$\rvir$ are shown in
    the {\it left panel} and (1.3-1.5)$\rvir$ in the {\it right
      panel}. The black lines show the CR electrons with a
    $\gam_\e=30$, blue lines $\gam_\e=200$, and green lines
    $\gam_\e=10^4$. The solid lines show the CR electrons with full
    cooling, i.e. for Coulomb, inverse Compton, and adiabatic
    losses. Dashed lines show the CR electrons without Coulomb and
    inverse Compton cooling while the dash-dotted lines show the CR
    electrons without adiabatic cooling. The dotted line show the
    injected CR electron distribution function, derived from the
    injected CR protons. For each energy, we normalize the CR electron
    distribution function with the CR electrons without
    cooling. Notice that the high energy part is built up during the
    last Gyr. This is explained by the inverse Compton losses that are
    larger at high energies, especially at high redshifts where the
    energy density of the CMB is much larger than
    today.  \label{fig:e_evol}}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
%  \includegraphics[width=0.49\columnwidth]{./figures/evolution.0.3Rv.eps}
%  \includegraphics[width=0.49\columnwidth]{./figures/evolution.1.4Rv.eps}
  \caption{Time evolution of the gas density and the cosmic ray (CR)
    cooling time. We show the time evolution of the electron number
    density and CR electron cooling time that are associated with the
    CR electrons which at redshift $z=0$ reside from (0.3-0.5)$\rvir$
    in the {\it left panel} and (1.3-1.5)$\rvir$ in the {\it right
      panel}. The red lines show the electron number densities, where
    the dashed line show the median and dash-dotted the mean. The
    solid lines show the cooling times
    $\tau_\rmn{cool}=\gam_\e/[b_\IC(\gam_\e)+b_\rmn{C}(\gam_\e)]$
    for $\gam_\e=30$ (black), $\gam_\e=200$ (blue), and
    $\gam_\e=10^4$ (green). Interesting the CR cooling at low
    energies of the particles that end up in the center is constant
    while is increasing monotonically for the particles that end up in
    the cluster periphery. The density of the particles that end up in
    the center is roughly constant with time while it is decreasing
    for the particles that end up in the outer part. This is due to
    the expansion of the universe that counteract the contraction of
    the particles in the cluster periphery. \label{fig:evol}}
\end{minipage}
\end{figure*}
